Low Complexity Adaptive Sonar Imaging

### 2018-10-11 LCAforActiveSonarImaging/document.tex
All compiling, but lots of work to be done
Testing change of frame letters

\title{Low Complexity Adaptive Sonar Imaging}

@@

This article is outlined as follows: In Section \ref{sec:beamforming} we offer a gentle introduction to beamforming. Then we move on to adaptive beamforming and describe the MVDR and LCA methods in \ref{sec:mvdr} and \ref{sec:lca}, respectively. We use LCA with the Kaiser window which we describe in \ref{sec:lca_kaiser_windows}, explain how we steer it in \ref{sec:lca_steering}, and by how much in \ref{sec:lca_steering_bounds}. Finally we add some remarks on using the trigonometric window instead of Kaiser in \ref{sec:lca_trigonometric}, and of of sampling considerations in \ref{sec:lca_oversampling}. Results and discussion are provided in Section \ref{sec:results_discussion}, where we study MVDR's windows to find suitable LCA ones in \ref{sec:results_lca_window_function}, assess the range of window types and steering in \ref{sec:results_window_parameters}, determine the number of windows needed in \ref{sec:results_database_size}, and add a note on computational complexity in \ref{sec:results_complexity}. Finally, Section \ref{sec:conclusion} provides a conclusion.

\section{Receive Beamforming}\label{sec:beamforming}

@@

The output $z[n]$ of a beamformer is defined as the weighted sum of all the delayed data samples:
%
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{eq:beamformer_output}
\end{align}
%
where $w_m$ is the weight factor assigned to channel $m$. A weight set is commonly called a window, apodization or taper function. When the window is real its response is symmetric. Applying a real window that trails off towards the edges creates a response with lower sidelobe levels and wider mainlobe, which translates into improved noise suppression at the cost of reduced resolution, respectively~\cite{Harris1978}.

@@

\section{MVDR}\label{sec:mvdr}

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\argmin{\w[n]}\, E\{\big|z[n]\big|^2\} &= \argmin{\w[n]}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

\ifPhdDoc
   \begin{figure}[tbp]\centering%
   \includegraphics[width=.8\linewidth]{gfx/fig1.eps}%
\else
   \begin{figure}[tbp]\centering%
   \ifPeerReview%
      \includegraphics[width=.5\linewidth]{gfx/buske1_online.eps}%
   \else%
      \includegraphics[width=\linewidth]{gfx/buske1_online.eps}%
\fi\fi%
\caption{Each Kaiser window is steered in the interval $\phi\in[0, \phi_{\mathrm{3\,dB}}(\beta)]$. The angle $\phi_{\mathrm{3\,dB}}(\beta)$ is the amount of steering needed for the steered amplitude response to have a \protect\minus{}3\,dB crossing that is exactly half that of the unsteered window. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.}\label{windows_steering}
\end{figure}

\ifPhdDoc
   \begin{figure}[tbp]\centering%
   \includegraphics[width=.7\linewidth]{gfx/fig2.png}%
\else
   \begin{figure}[tbp]\centering%
   \ifPeerReview%
      \includegraphics[width=.5\linewidth]{gfx/buske2_online.png}%
   \else%
      \includegraphics[width=\linewidth]{gfx/buske2_online.png}%
\fi\fi%
\caption{The submerged object used to test beamforming resolution is a 1\,m by 1\,m test cross attached to an anchor with a diameter of approximately 13\,cm. Source image curtesy of Bundeswehr Technical Center for Ships and Naval Weapons, Maritime Technology and Research (WTD 71).}\label{cross}
\end{figure}

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig3_phd.pdf}%
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\\[.5\baselineskip]
   \textbf{Left:} MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.\\[.5\baselineskip]
   \textbf{Right:} MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.}\label{mvdr_selected_windows}
\else
\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

   \begin{figure*}[t]\centering%
   \includegraphics[width=\linewidth]{gfx/buske3_online.pdf}%
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
   \emph{Left:}\hfill
   \parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.}\protect\\\hspace{\textwidth}
   \emph{Right:}\hfill
   \parbox[t]{.95\linewidth}{MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.} }\label{mvdr_selected_windows}
\fi%

@@

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig4.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \\[.5\baselineskip]
   \textbf{Upper bound $\beta$:} Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.
   \\[.5\baselineskip]
   \textbf{Upper bound $\phi$:} As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution.
   }
\else
   \begin{figure*}[t]%
   \includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \newline
   \emph{Upper bound $\beta$:}\hfill
   \parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
   \emph{Upper bound $\phi$:}\hfill
   \parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
   }
\fi%
\label{oversampling_mosaic_bounds}
\end{figure*}

@@

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/fig5a.pdf}%%
   \label{oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/fig5b.pdf}%
   \label{oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{fig_sim}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/buske5a_online.pdf}%%
   \label{oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/buske5b_online.pdf}%
   \label{oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{fig_sim}
   \end{figure*}
\fi

\ifPhdDoc
   \begin{figure*}[tb]\centering%
   \includegraphics[width=\linewidth]{gfx/fig6.pdf}%
\else
   \begin{figure*}[t]%\centering%
   \includegraphics[width=\textwidth]{gfx/buske6_online.pdf}%
\fi
\caption{\emph{LCA images using different window database sizes}. The window boundaries are the same. The upper left image is identical to a rectangularly weighted DAS. Observe that finer sampling of the $\beta$-range improves sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. However, using more than  $N_\beta=2$ window types and $N_\phi=3$ steering angles makes minimal difference.}\label{oversampling_mosaic}
\end{figure*}

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/fig7a.pdf}%
   \label{beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/fig7b.pdf}%%
   \label{beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{image_quality}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/buske7a_online.pdf}%
   \label{beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/buske7b_online.pdf}%%
   \label{beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{image_quality}
   \end{figure*}
\fi

\section{LCA}\label{sec:lca}

A much less complex alternative to MVDR is the low complexity adaptive (LCA) beamformer. It iterates through a set of $P$ windows and selects the window $p$ that best fulfills the minimum variance criterion:
%
\begin{align}
\argmin{p}\, E\{\big|z_p[n]\big|^2\} = \argmin{p}\, E\big\{\big|\w_{p}\H\x[n]\big|^2\big\}\nn
\qquad\text{subject to}\qquad \w\H\a_\phi = 1.\label{eq:lca_definition}
\end{align}\label{lca_criterion}
%
Note how closely LCA is related to the MVDR definition in \eq{eq:mvdr_definition}. The optimization criterion and constraint is the same, but LCA has a finite and discrete solution space for the weights. As will be demonstrated in Section \ref{sec:results_lca_window_function}, LCA performs similarly to MVDR because a robustified MVDR implementation seems to be constrained to window functions similar to the ones we let LCA choose from.

@@

If the pixels are correlated laterally, we can include a weighted combination of these to improve the variance estimation:
%
\begin{align}
s_z^2 = \frac{1}{N_x N_k} \sumb{\nmat{\scriptstyle x'=\\[-2mm]\scriptstyle x-X}}{x+X} \sumb{\nmat{\scriptstyle n'=\\[-2mm]\scriptstyle n-K}}{n+K} \omega[x',n']\big| z[x',n'] \big|^2
\end{align}
%
where $N_x = 2X+1$ is the number of azimuth lines to average over, and $\omega[x',n']$ is a normalized 2 dimensional weight function. Since we oversample slightly laterally when we delay each pixel (see Section \ref{sec:lca_oversampling}), we thought the lateral correlation to be sufficient to benefit from this. However, we observed no visual improvement from applying this technique, and decided not to use it.
%

\subsection{Window function: Kaiser}\label{sec:lca_kaiser_windows}

>>>>

% !TeX root = III_lca
\title{Low Complexity Adaptive Sonar Imaging}

@@

This article is outlined as follows: In Section \ref{III_sec:beamforming} we offer a gentle introduction to beamforming. Then we move on to adaptive beamforming and describe the MVDR and LCA methods in \ref{III_sec:mvdr} and \ref{III_sec:lca}, respectively. We use LCA with the Kaiser window which we describe in \ref{III_sec:lca_kaiser_windows}, explain how we steer it in \ref{III_sec:lca_steering}, and by how much in \ref{III_sec:lca_steering_bounds}. Finally we add some remarks on using the trigonometric window instead of Kaiser in \ref{III_sec:lca_trigonometric}, and of of sampling considerations in \ref{III_sec:lca_oversampling}. Results and discussion are provided in Section \ref{III_sec:results_discussion}, where we study MVDR's windows to find suitable LCA ones in \ref{III_sec:results_lca_window_function}, assess the range of window types and steering in \ref{III_sec:results_window_parameters}, determine the number of windows needed in \ref{III_sec:results_database_size}, and add a note on computational complexity in \ref{III_sec:results_complexity}. Finally, Section \ref{III_sec:conclusion} provides a conclusion.

\section{Receive Beamforming}\label{III_sec:beamforming}

@@

The output $z[n]$ of a beamformer is defined as the weighted sum of all the delayed data samples:
%
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{III_eq:beamformer_output}
\end{align}
%
where $w_m$ is the weight factor assigned to channel $m$. A weight set is commonly called a window, apodization or taper function. When the window is real its response is symmetric. Applying a real window that trails off towards the edges creates a response with lower sidelobe levels and wider mainlobe, which translates into improved noise suppression at the cost of reduced resolution, respectively~\cite{Harris1978}.

@@

\section{MVDR}\label{III_sec:mvdr}

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\argmin{\w[n]}\, E\{\big|z[n]\big|^2\} &= \argmin{\w[n]}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{III_eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{III_eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

\ifPhdDoc
   \begin{figure}[tbp]\centering%
   \includegraphics[width=.8\linewidth]{gfx/fig1.eps}%
\else
   \begin{figure}[tbp]\centering%
   \ifPeerReview%
      \includegraphics[width=.5\linewidth]{gfx/buske1_online.eps}%
   \else%
      \includegraphics[width=\linewidth]{gfx/buske1_online.eps}%
\fi\fi%
\caption{Each Kaiser window is steered in the interval $\phi\in[0, \phi_{\mathrm{3\,dB}}(\beta)]$. The angle $\phi_{\mathrm{3\,dB}}(\beta)$ is the amount of steering needed for the steered amplitude response to have a \protect\minus{}3\,dB crossing that is exactly half that of the unsteered window. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.}\label{III_windows_steering}
\end{figure}

\ifPhdDoc
   \begin{figure}[tbp]\centering%
   \includegraphics[width=.7\linewidth]{gfx/fig2.png}%
\else
   \begin{figure}[tbp]\centering%
   \ifPeerReview%
      \includegraphics[width=.5\linewidth]{gfx/buske2_online.png}%
   \else%
      \includegraphics[width=\linewidth]{gfx/buske2_online.png}%
\fi\fi%
\caption{The submerged object used to test beamforming resolution is a 1\,m by 1\,m test cross attached to an anchor with a diameter of approximately 13\,cm. Source image curtesy of Bundeswehr Technical Center for Ships and Naval Weapons, Maritime Technology and Research (WTD 71).}\label{III_cross}
\end{figure}

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig3_phd.pdf}%
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\\[.5\baselineskip]
   \textbf{Left:} MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.\\[.5\baselineskip]
   \textbf{Right:} MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.}\label{III_mvdr_selected_windows}
\else
\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

   \begin{figure*}[t]\centering%
   \includegraphics[width=\linewidth]{gfx/buske3_online.pdf}%
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
   \emph{Left:}\hfill
   \parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.}\protect\\\hspace{\textwidth}
   \emph{Right:}\hfill
   \parbox[t]{.95\linewidth}{MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.} }\label{III_mvdr_selected_windows}
\fi%

@@

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig4.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \\[.5\baselineskip]
   \textbf{Upper bound $\beta$:} Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.
   \\[.5\baselineskip]
   \textbf{Upper bound $\phi$:} As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution.
   }
\else
   \begin{figure*}[t]%
   \includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \newline
   \emph{Upper bound $\beta$:}\hfill
   \parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
   \emph{Upper bound $\phi$:}\hfill
   \parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
   }
\fi%
\label{III_oversampling_mosaic_bounds}
\end{figure*}

@@

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/fig5a.pdf}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/fig5b_phd.pdf}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/buske5a_online.pdf}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/buske5b_online.pdf}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\fi

\ifPhdDoc
   \begin{figure*}[tb]\centering%
   \includegraphics[width=\linewidth]{gfx/fig6.pdf}%
\else
   \begin{figure*}[t]%\centering%
   \includegraphics[width=\textwidth]{gfx/buske6_online.pdf}%
\fi
\caption{\emph{LCA images using different window database sizes}. The window boundaries are the same. The upper left image is identical to a rectangularly weighted DAS. Observe that finer sampling of the $\beta$-range improves sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. However, using more than  $N_\beta=2$ window types and $N_\phi=3$ steering angles makes minimal difference.}\label{III_oversampling_mosaic}
\end{figure*}

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/fig7a.pdf}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/fig7b_phd.pdf}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/buske7a_online.pdf}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/buske7b_online.pdf}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\fi

\section{LCA}\label{III_sec:lca}

A much less complex alternative to MVDR is the low complexity adaptive (LCA) beamformer. It iterates through a set of $P$ windows and selects the window $p$ that best fulfills the minimum variance criterion:
%
\begin{align}
\argmin{p}\, E\{\big|z_p[n]\big|^2\} = \argmin{p}\, E\big\{\big|\w_{p}\H\x[n]\big|^2\big\}\nn
\qquad\text{subject to}\qquad \w\H\a_\phi = 1.\label{III_eq:lca_definition}
\end{align}\label{III_lca_criterion}
%
Note how closely LCA is related to the MVDR definition in \eq{III_eq:mvdr_definition}. The optimization criterion and constraint is the same, but LCA has a finite and discrete solution space for the weights. As will be demonstrated in Section \ref{III_sec:results_lca_window_function}, LCA performs similarly to MVDR because a robustified MVDR implementation seems to be constrained to window functions similar to the ones we let LCA choose from.

@@

If the pixels are correlated laterally, we can include a weighted combination of these to improve the variance estimation:
%
\begin{align}
s_z^2 = \frac{1}{N_x N_k} \sumb{\nmat{\scriptstyle x'=\\[-2mm]\scriptstyle x-X}}{x+X} \sumb{\nmat{\scriptstyle n'=\\[-2mm]\scriptstyle n-K}}{n+K} \omega[x',n']\big| z[x',n'] \big|^2
\end{align}
%
where $N_x = 2X+1$ is the number of azimuth lines to average over, and $\omega[x',n']$ is a normalized 2 dimensional weight function. Since we oversample slightly laterally when we delay each pixel (see Section \ref{III_sec:lca_oversampling}), we thought the lateral correlation to be sufficient to benefit from this. However, we observed no visual improvement from applying this technique, and decided not to use it.
%

\subsection{Window function: Kaiser}\label{III_sec:lca_kaiser_windows}

As will be demonstrated in upcoming sections, the LCA beamformer works very well with windows generated from the Kaiser-Bessel function. We will express it in vector form as:
%
\begin{align}
\f_\beta = \bmat{
f_0(\beta) \\
\vdots\\
f_{M-1}(\beta)
}\label{III_eq:kaiser_window_vector}
\end{align}
%
where
%
\begin{align}
f_m(\beta) = \frac{I_0\left(\pi\beta\sqrt{1-\left(\frac{2m}{M-1}-1\right)^2}\right)}{I_0(\pi\beta)}\label{III_eq:kaiser_window_element}
\end{align}
%
and $I_0$ is the zeroth order modified Bessel function of the first kind:
%
\begin{align}
I_0(x) = \sumb{a=0}{\infty} \left[ \frac{\left(\frac{x}{2}\right)^a}{a!} \right]^2.\label{III_eq:modified_bessel_first_kind}
\end{align}
%
The Kaiser-Bessel window is near optimal in the sense of having its peak energy concentration around $\theta=0^\circ$, for a given space-bandwidth product related to the Kaiser parameter $\beta$ as:
%
\begin{align}
\beta = \frac{TB}{2},
\end{align}
%
where $T$ in our case is the spatial extent of the window and $B$ is its bandwidth. Adjusting $\beta$ changes the trade-off between mainlobe width and sidelobe level. When $\beta=0$ the window becomes rectangular, while at large values ($\beta>5$) the window converges to a Gaussian both in time and frequency. This class of windows is generally considered well suited for separating closely spaced sources with amplitudes of a high dynamic range~\cite{Harris1978}, they are easy to make, and they are optimal for any value of $\beta$.

\subsection{Steering}\label{III_sec:lca_steering}
